Sequential rank agreement methods for comparison of ranked lists

### 2015-07-08 latex/paper.Rnw
Andreas' application - part 1

\subsection{Comparing results across different method}

<<>>=
#library(changepoint)
#cp1 <- cpt.mean(res, method="BinSeg")
#cp2 <- cpt.mean(res2, method="BinSeg")
#cp1@cpts
#cp1@param.est
#cp2@cpts
#cp2@param.est
@

\subsection{Stability of results}

\subsection{Evaluating results from top-$k$ lists}

@@

\subsection{Changepoint analysis}

>>>>

We now consider an application of the sequential rank agreement to two
data sets consisting of MALDI-TOF (Matrix-Assisted Laser
Desorption/Ionization Time Of Flight) mass spectra obtained from blood
samples from patients with either benign or malignant ovarian
tumors. The data sets are subsamples of the Danish MALOVA and DACOVA
study populations.

The MALOVA study is a multidisciplinary Danish study on ovarian cancer
\cite{Hogdall:2004:Cancer:15160342} where all Danish women diagnosed
with an ovarian tumor and referred for surgery from the participating
departments of gynecology were enrolled continuously from December
1994 to May 1999. We use a random subsample of $119$ patients with a
total of $58$ patients with malignant ovarian cancers as cases and
$61$ patients with benign ovarian tumors as controls. The DACOVA study
is similarly a multidisciplinary Danish study on ovarian cancer
covering about $66\%$ of the population in Denmark
\cite{bertelsen1991protocol}. It was attempted to continuous enroll
all patients referred for surgery with an ovarian tumor clinically
suspected to be cancer during the period 1984 to 1990. We use a random
subsample of $113$ patients with a total of $54$ malignant ovarian
cancers and $59$ benign ovarian tumors/gynecologic disorders.

Proteomic analysis was performed by MALDI-TOF mass spectrometry with a
Bruker Ultraflex in positive, linear ion mode in the mass-to-charge
range $800$ to $20000$ Dalton by an average of $500$ laser
shots. Calibration of the instrument was performed using a protein
standard mixture. Each spectrum consists of $49642$ samples over the
mass-to-charge range which we down-sample on an equidistant grid of
5000 points by linear interpolation. We further preprocess each
spectrum by first removing the slow-varying baseline intensities with
the SNIP algorithm \cite{ryan1988snip} followed by a normalization
with respect to the total ion counts. Finally, we standardize the 5000
predictors to have column-wise zero mean and unit variance.

We use these two data sets to illustrate how the sequential rank
agreement can be applied in two different scenarios: i) assessing the
agreement of ranking the importance of the predictors using different
statistical classification methods and ii) assessing the agreement of
rankings the predictions (the probability of the classifying a tumor
as malignant) subject to perturbations in data. In the latter case we
further illustrate the discrepancy between the ranking stability of
different statistical methods through a simulation study.

\subsection{Agreement of predictor ranking}

#includegraphics marginalVsModel.pdf

\subsection{Agreement of predictions}

#includegraphics thomas.pdf

#includegraphics simulationTest2.pdf

@@

Changepoint analysis
